---
title: "Prediction Assignment"
author: "Eunice Ang"
date: "September 30, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

##Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

##a) Preparing the environment
Firstly, we load all the libraries that are required in our analysis.
```{r libraries}
rm(list=ls())                # free up memory for the download of the data sets
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(55)
```

##b) Data Loading and Cleaning
Next, we load the dataset from the URLs provided above. The training dataset is split into a Training set (70% of the data) for the modeling process and a Test set (with the remaining 30%) for the validations.

```{r data}
# Create variables for the URLs for the download of data
Url_Train <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Url_Test <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# Download the datasets
training <- read.csv(url(Url_Train))
testing  <- read.csv(url(Url_Test))

# Split the training dataset into test set and training set
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
```

```{r}
dim(TestSet)
```

Both created datasets have 160 variables. Now, we clean the data by removing NA values. The Near Zero variance (NZV) variables are also removed and the ID variables as well.

```{r clean}
# remove variables with Nearly Zero Variance
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)
```

```{r}
dim(TestSet)
```

```{r}
# remove variables that are mostly NA
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
```

```{r}
dim(TestSet)
```

```{r}
# remove identification only variables (columns 1 to 5)
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
```

```{r}
dim(TestSet)
```
After cleaning, the number of variables for the analysis has been reduced to 54 only.

##c) Prediction Model Building
```{r}
set.seed(55)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel
```

```{r}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
